---
title: "p8105_hw2_sl4660"
author: "Sibei Liu"
date: "2019/10/1"
output: github_document
---
## Problem 1
```{r,message=FALSE}

library(tidyverse)
library(tidyr)
library(dplyr)
library(readxl)
```

```{r,warning=FALSE}
trash_wheel=read_excel("./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx",sheet=1,range="A2:N408") %>% 
   janitor::clean_names()
 trash_wheel_md=drop_na(trash_wheel,dumpster) %>% 
   mutate(sports_balls=as.integer(round(sports_balls)))
 trash_wheel_md
```

```{r}
pre_2018=read_excel("./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx",sheet=5,skip=1) %>% 
  janitor::clean_names() %>% 
  drop_na(total,month) %>% 
  mutate(year="2018")


pre_2017=read_excel("./data/HealthyHarborWaterWheelTotals2018-7-28.xlsx",sheet=7,skip=1) %>% 
 janitor::clean_names() %>% 
  drop_na(total,month) %>% 
  mutate(year="2017")
 
binded_data=bind_rows(pre_2017,pre_2018) %>% 
  mutate(
    month=recode(month,`1`=month.name[1],`2`=month.name[2],`3`=month.name[3],
                 `4`=month.name[4],`5`=month.name[5],`6`=month.name[6],
                 `7`=month.name[7],`8`=month.name[8],`9`=month.name[9],
                 `10`=month.name[10],`11`=month.name[11],`12`=month.name[12])
  )
binded_data
```
**Descriptions**

In cleaned dataset trash_wheel,there are ``r nrow(trash_wheel_md)`` rows and ``r ncol(trash_wheel_md)`` columns. I think the key variable to locate the row we want is dumpster.Because dumpter has the unique number in each rows. The tables consisit of three different information. First is the key variable:dumspter number. Second is the date collection. Third is the type and amount of litters including variables weight_tons,volume_cubic_yard,plastic_bottles,polystyrene,cigarette_butts,glass_bottles,grocery_bags,chip_bags,sports_balls

In cleaned and combined dataset presipitation,there are ``r nrow(binded_data)`` rows and ``r ncol(binded_data)`` columns. I think the key variables to locate the row we want are year and month.Because only according to the year and month,could we locate the detailed presipitation amount.

The total precipitation in 2018 is ``r sum(select(pre_2018,total))``.
The median number of sports balls in a dumpster in 2017 is 
`r median(pull(filter(trash_wheel_md,year==2017),sports_balls))`

## Problem 2

# Frist
```{r}
pols_month=read_csv("./data/fivethirtyeight_datasets/pols-month.csv") %>% 
 separate(col=mon,into=c("year","month","day"),sep="/") 
  pols_month2=mutate(pols_month,
   year=as.integer(year),
   day=as.integer(day),
   month=as.integer(month),
   month=recode(month,`1`=month.name[1],`2`=month.name[2],`3`=month.name[3],
                 `4`=month.name[4],`5`=month.name[5],`6`=month.name[6],
                 `7`=month.name[7],`8`=month.name[8],`9`=month.name[9],
                 `10`=month.name[10],`11`=month.name[11],`12`=month.name[12]),
   persident=ifelse(pull(pols_month,prez_gop)==0, "dem","gop")) 
  pols_month3=select(pols_month2,-prez_dem,-prez_gop,-day)
  pols_month3
```

# Second
```{r}
snp=read_csv("./data/fivethirtyeight_datasets/snp.csv") %>% 
 separate(col=date,into=c("month","day","year"),sep="/") %>% 
 mutate(
   day=as.integer(day),
   year=as.integer(year),
   month=as.integer(month)) %>% 
   arrange(year,month) %>% 
mutate(month=recode(month,`1`=month.name[1],`2`=month.name[2],`3`=month.name[3],
                 `4`=month.name[4],`5`=month.name[5],`6`=month.name[6],
                 `7`=month.name[7],`8`=month.name[8],`9`=month.name[9],
                 `10`=month.name[10],`11`=month.name[11],`12`=month.name[12])) %>% 
select(-day) %>% 
  select(year,month,close)
snp
```

# Third
```{r}
 unemploy=read_csv("./data/fivethirtyeight_datasets/unemployment.csv")
unemploy_tidy_data = 
  pivot_longer(
    unemploy, 
    Jan:Dec,
    names_to = "month", 
    values_to = "percentage") 
 unemploy2=janitor::clean_names(unemploy_tidy_data)
   unemploy3=mutate(unemploy2,
   month=recode(month,"Jan"=month.name[1],"Feb"=month.name[2],"Mar"=month.name[3],
                 "Apr"=month.name[4],"May"=month.name[5],"Jun"=month.name[6],
                 "Jul"=month.name[7],"Aug"=month.name[8],"Sep"=month.name[9],
                 "Oct"=month.name[10],"Nov"=month.name[11],"Dec"=month.name[12])
   ) 
   unemploy3
```

## Mergeing
```{r}
total1=left_join(pols_month3,snp)
total2=left_join(total1,unemploy3)
total2
```
## Description of Problem 2

In cleanded dataset of pols_month, there are `r nrow(pols_month3)` rows and `r ncol(pols_month3)` columns. The range of year is `r min(pull(pols_month3,year))` to `r max(pull(pols_month3,year))`.The key variables in this dataset are year and month. The pols_month table represent the number of national politicians who are democratic or republican during this time.


In cleanded dataset of snp, there are `r nrow(snp)` rows and `r ncol(snp)` columns. The range of year is `r min(pull(snp,year))` to `r max(pull(snp,year))`.The key variables in this dataset are year and month.

In cleanded dataset of unemplyment, there are `r nrow(unemploy3)` rows and `r ncol(unemploy3)` columns. The range of year is `r min(pull(unemploy3,year))` to `r max(pull(unemploy3,year))`.The key variables in this dataset are year and month.

In cleanded dataset of merged dataset, there are `r nrow(total2)` rows and `r ncol(total2)` columns. The range of year is `r min(pull(total2,year))` to `r max(pull(total2,year))`.The key variables in this dataset are year and month.

# Problem3
```{r}
pop_name=read_csv("./data/Popular_Baby_Names.csv") %>% 
  janitor::clean_names() %>% 
    mutate(gender=str_to_lower(gender),
           ethnicity=str_to_lower(ethnicity),
           childs_first_name=str_to_lower(childs_first_name),
          ethnicity=recode(ethnicity,"asian and paci"="asian and pacific islander","black non hisp"="black non hispanic","white non hisp"="white non hispanic"))
pop_name2=distinct(pop_name)
pop_name2
```
```{r}
olivia=filter(pop_name2,gender=="female"&childs_first_name=="olivia") %>% 
select(year_of_birth,ethnicity,rank) %>% 
pivot_wider(
  names_from = "year_of_birth", 
  values_from = "rank")
olivia
```

```{r}
pop_male_name=filter(pop_name2,gender=="male"&rank==1) %>% 
select(year_of_birth,ethnicity,childs_first_name) %>% 
pivot_wider(
  names_from = "year_of_birth", 
  values_from = "childs_first_name")
pop_male_name
```