---
title: "p8105_hw2_sl4660"
author: "Sibei Liu"
date: "2019/10/1"
output: github_document
---
## Problem 1
```{r,message=FALSE}

library(tidyverse)
library(tidyr)
library(dplyr)
library(readxl)
```

```{r,warning=FALSE}
trash_wheel=read_excel("./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx",sheet=1,range="A2:N408") %>% 
   janitor::clean_names()
 trash_wheel_md=drop_na(trash_wheel,dumpster) %>% 
   mutate(sports_balls=as.integer(round(sports_balls)))
 trash_wheel_md
```

```{r}
pre_2018=read_excel("./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx",sheet=5,skip=1) %>% 
  janitor::clean_names() %>% 
  drop_na(total,month) %>% 
  mutate(year="2018")


pre_2017=read_excel("./data/HealthyHarborWaterWheelTotals2018-7-28.xlsx",sheet=7,skip=1) %>% 
 janitor::clean_names() %>% 
  drop_na(total,month) %>% 
  mutate(year="2017")
 
binded_data=bind_rows(pre_2017,pre_2018) %>% 
  mutate(
    month=recode(month,`1`=month.name[1],`2`=month.name[2],`3`=month.name[3],
                 `4`=month.name[4],`5`=month.name[5],`6`=month.name[6],
                 `7`=month.name[7],`8`=month.name[8],`9`=month.name[9],
                 `10`=month.name[10],`11`=month.name[11],`12`=month.name[12])
  )
binded_data